---
title: "Combined Digital Images"
output:
  pdf_document: default
  
---


```{r message=FALSE, warning=FALSE, include=FALSE}
setwd("~/Documents/OneDrive - University Of Cambridge/Documents/PhD/Neoadjuvant/rscripts/")
library(readr)
library(dplyr)
library(knitr)
library(kableExtra)

combined_ffpe <- read_csv("~/Documents/OneDrive - University Of Cambridge/Documents/PhD/Neoadjuvant/Scanned Images/Combined_ffpe.csv")
```


The trials included in this dataset are:
 - Monet
 - Persephone
 - TransNeo
 - Partner
 
Each trial has the following:
```{r eval=FALSE, message=FALSE, warning=FALSE, include=FALSE, paged.print=FALSE}
monet <- combined_ffpe[combined_ffpe$Trial == "Monet", ]
persephone <-  combined_ffpe[combined_ffpe$Trial  == "Persephone", ]
transneo <- combined_ffpe[combined_ffpe$Trial  == "TransNeo", ]
partner <- combined_ffpe[combined_ffpe$Trial  == "Partner", ]

trial_cases <- as.data.frame(cbind(c("Monet","Persephone", "TransNeo", "Partner"), c(length(unique(monet$Trial_ID)), length(unique(persephone$Trial_ID)), length(unique(transneo$Trial_ID)), length(unique(partner$Trial_ID)))))
colnames(trial_cases) <- c("Trial", "Number of Patients")

kable(trial_cases, align = "c") %>% 
  kable_styling(full_width = F, font_size =11)
```

 

We have slides that have H&E and other stains, from both biopsy and surgical procedures.  
```{r echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}

biopsy <- combined_ffpe[combined_ffpe$Procedure=="Biopsy", ]
surgical <- combined_ffpe[combined_ffpe$Procedure=="Surgical", ]

trial_biopsy <- as.data.frame.matrix(table(biopsy$Trial, biopsy$Stain=="H&E"))
trial_biopsy <- trial_biopsy[ ,c(2,1)]
colnames(trial_biopsy) <- c("H and E", "Others")
trial_biopsy$Total <- rowSums(trial_biopsy, dims=1)

trial_surgical <- as.data.frame.matrix(table(surgical$Trial, surgical$Stain=="H&E"))
trial_surgical <- trial_surgical[ ,c(2,1)]
colnames(trial_surgical) <- c("H and E", "Others")
trial_surgical$Total <- rowSums(trial_surgical, dims=1)

trials <- cbind.data.frame(trial_biopsy, trial_surgical)
trials$Overall <- trial_biopsy$Total + trial_surgical$Total
trials <- rbind(trials, colSums(trials, dims=1))
rownames(trials)[5] <- "Total"
trials$'Number of Patients' <- c(26, 32, 33, 180, 271)
kable(trials, escape=F, align="c") %>%
  kable_styling(full_width = F, "bordered", font_size =11) %>% 
  add_header_above(c("Trial" = 1, "Biopsy" = 3, "Surgical"=3, " " = 1, "  " = 1 ))

```

Of these, we have slides of breast tissue and lymph node tissue.  Some fibrousfatty tissue is also included as they were taken with lymph nodes

```{r echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
trial_breast <- combined_ffpe[combined_ffpe$Tissue == "Breast", ]
trial_LN <- combined_ffpe[combined_ffpe$Tissue == "LN", ]

trial_tissue <- as.data.frame.matrix(table(combined_ffpe$Trial, combined_ffpe$Tissue))
trial_tissue <- trial_tissue[, c(2, 5, 1, 3, 4, 6)]

kable(trial_tissue, align = "c") %>%
  kable_styling(full_width = F, font_size =11)
```

Amongst the breast slides, the number of slides that include the following features are:

```{r echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
invasive <- as.data.frame.matrix(table(trial_breast$Trial, trial_breast$Tumour))
InSitu <- as.data.frame.matrix(table(trial_breast$Trial, trial_breast$InSitu))
LVI <- as.data.frame.matrix(table(trial_breast$Trial, trial_breast$LVI))
LVI$Trials <- rownames(LVI)
tumour <- merge(invasive, InSitu,  by="row.names")
colnames(tumour)[1] <- "Trials"
tumour <- merge(tumour, LVI, by = "Trials")



kable(tumour, align="c") %>%
  kable_styling(full_width = F, font_size =11)



```


The number of lymph node per slides varies.  

```{r echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
LNperslide <- table(trial_LN$Trial,  trial_LN$TotalLN)
kable(LNperslide) %>% 
  kable_styling(full_width = F, font_size =11)
```



