---
title: "Neighbours and Distances"
output: pdf_document
---

The analysis is performed using the SVM version of cell extractor, with normal, cancer, lymphocyte, stroma as cell categories.  

Aim:
1. Look at the composition of a cell's neighbours.  how many are cancer, how many are stroma etc
2. Look at the distance between cell types and see whether one are closer to another


the cell key is as following:
1 is tumour, 2 lymph, 3 stroma, 4 normal


Firstly we look at a cell's neighbours

```{r}
library(reshape2)
library(ggplot2)
library(readr)

```



```{r}
neighbours_605012_lymph <- read_csv("~/Documents/OneDrive - University Of Cambridge/Documents/PhD/Neoadjuvant/Thomas/Cell_clustering/detail/clustering_detail_605012_lymphocyte_cellids_nearest50.csv", col_names = FALSE)

neighbours_605012_lymph$ID <- paste("Lymph",c(1:nrow(neighbours_605012_lymph)))
rownames(neighbours_605012_lymph) <- neighbours_605012_lymph$ID

neighbours_605012_lymph$Tumour <- rowSums(neighbours_605012_lymph==1)
neighbours_605012_lymph$Lymph <- rowSums(neighbours_605012_lymph==2)
neighbours_605012_lymph$Stroma <- rowSums(neighbours_605012_lymph==3)
neighbours_605012_lymph$Normal <- rowSums(neighbours_605012_lymph==4)

neighbours_605012_lymph_by_position <- data.frame(cbind(colSums(neighbours_605012_lymph==1)+colSums(neighbours_605012_lymph==4), colSums(neighbours_605012_lymph==2), colSums(neighbours_605012_lymph==3)))
colnames(neighbours_605012_lymph_by_position) <- c("Tumour", "Lymphocytes", "Stroma")
neighbours_605012_lymph_by_position$Position <- rownames(neighbours_605012_lymph_by_position)
neighbours_605012_lymph_by_position$Position[1:9] <- c("X01", "X02", "X03", "X04", "X05", "X06", "X07", "X08", "X09")
neighbours_605012_lymph_by_position_melted <- melt(neighbours_605012_lymph_by_position[1:50,])

ggplot(neighbours_605012_lymph_by_position_melted, aes(x=Position, y=value/nrow(neighbours_605012_lymph), group=variable, colour=variable))+
  geom_line(size=2)+
  theme_bw(base_size = 18)+ 
  scale_color_discrete("Neighbouring Cell Type")+
  labs(title="Lymphocytes neighbour cell type")+
  geom_hline(yintercept = c(0.081928864,	0.03723156,	0.880839577), colour=c("#F8766D", "#00BA38", "#619CFF" ), linetype="dashed")+ 
  theme(axis.text.x = element_text(angle = 90, hjust = 1))+
  labs(y="Percentage")




```



```{r}
neighbours_647364_lymph <- read_csv("~/Documents/OneDrive - University Of Cambridge/Documents/PhD/Neoadjuvant/Thomas/Cell_clustering/detail/clustering_detail_647364_lymphocyte_cellids_nearest50.csv", col_names = FALSE)

neighbours_647364_lymph$ID <- paste("Lymph",c(1:nrow(neighbours_647364_lymph)))
rownames(neighbours_647364_lymph) <- neighbours_647364_lymph$ID

neighbours_647364_lymph$Tumour <- rowSums(neighbours_647364_lymph==1)
neighbours_647364_lymph$Lymph <- rowSums(neighbours_647364_lymph==2)
neighbours_647364_lymph$Stroma <- rowSums(neighbours_647364_lymph==3)
neighbours_647364_lymph$Normal <- rowSums(neighbours_647364_lymph==4)

neighbours_647364_lymph_by_position <- data.frame(cbind(colSums(neighbours_647364_lymph==1), colSums(neighbours_647364_lymph==2), colSums(neighbours_647364_lymph==3), colSums(neighbours_647364_lymph==4)))
colnames(neighbours_647364_lymph_by_position) <- c("Tumour", "Lymphocytes", "Stroma", "Normal")
neighbours_647364_lymph_by_position$Position <- rownames(neighbours_647364_lymph_by_position)
neighbours_647364_lymph_by_position$Position[1:9] <- c("X01", "X02", "X03", "X04", "X05", "X06", "X07", "X08", "X09")
neighbours_647364_lymph_by_position_melted <- melt(neighbours_647364_lymph_by_position[1:50,])

ggplot(neighbours_647364_lymph_by_position_melted, aes(x=Position, y=value, group=variable, colour=variable))+
  geom_line(size=2)+
  theme_bw()+ 
  scale_color_discrete("Neighbouring Cell Type")+
  labs(title="Lymphocytes neighbour cell type")


```




```{r}
neighbours_648121_lymph <- read_csv("~/Documents/OneDrive - University Of Cambridge/Documents/PhD/Neoadjuvant/Thomas/Cell_clustering/detail/clustering_detail_648121_lymphocyte_cellids_nearest50.csv", col_names = FALSE)

neighbours_648121_lymph$ID <- paste("Lymph",c(1:nrow(neighbours_648121_lymph)))
rownames(neighbours_648121_lymph) <- neighbours_648121_lymph$ID

neighbours_648121_lymph$Tumour <- rowSums(neighbours_648121_lymph==1)
neighbours_648121_lymph$Lymph <- rowSums(neighbours_648121_lymph==2)
neighbours_648121_lymph$Stroma <- rowSums(neighbours_648121_lymph==3)
neighbours_648121_lymph$Normal <- rowSums(neighbours_648121_lymph==4)

neighbours_648121_lymph_by_position <- data.frame(cbind(colSums(neighbours_648121_lymph==1)+colSums(neighbours_648121_lymph==4), colSums(neighbours_648121_lymph==2), colSums(neighbours_648121_lymph==3)))
#neighbours_648121_lymph_by_position$Tumour <- sum(neighbours_648121_lymph_by_position$Tumour, neighbours_648121_lymph_by_position$Normal)

colnames(neighbours_648121_lymph_by_position) <- c("Tumour", "Lymphocytes", "Stroma")
neighbours_648121_lymph_by_position$Position <- rownames(neighbours_648121_lymph_by_position)
neighbours_648121_lymph_by_position$Position[1:9] <- c("X01", "X02", "X03", "X04", "X05", "X06", "X07", "X08", "X09")
neighbours_648121_lymph_by_position_melted <- melt(neighbours_648121_lymph_by_position[1:50,])

ggplot(neighbours_648121_lymph_by_position_melted, aes(x=Position, y=value/nrow(neighbours_648121_lymph), group=variable, colour=variable))+
  geom_line(size=2)+
  theme_bw(base_size = 18)+ 
  scale_color_discrete("Neighbouring Cell Type")+
  labs(title="Lymphocytes neighbour cell type")+
  geom_hline(yintercept = c(0.409369812,	0.079632456,	0.510997732), colour=c("#F8766D", "#00BA38", "#619CFF" ), linetype="dashed")+ 
  theme(axis.text.x = element_text(angle = 90, hjust = 1))+
  labs(y="Percentage")


neighbours_648121_lymph_by_position_melted_1to10 <- melt(neighbours_648121_lymph_by_position[1:10,])
ggplot(neighbours_648121_lymph_by_position_melted_1to10, aes(x=Position, y=value, group=variable, colour=variable))+
  geom_line(size=2)+
  theme_bw()+ 
  scale_color_discrete("Neighbouring Cell Type")+
  labs(title="Lymphocytes neighbour cell type")


```



What about neighbours of tumour cells?

```{r}
neighbours_648121_tumour <- read_csv("~/Documents/OneDrive - University Of Cambridge/Documents/PhD/Neoadjuvant/Thomas/Cell_clustering/detail/clustering_detail_648121_tumour_cellids_nearest50.csv", col_names = FALSE)

neighbours_648121_tumour$ID <- paste("tumour",c(1:nrow(neighbours_648121_tumour)))
rownames(neighbours_648121_tumour) <- neighbours_648121_tumour$ID

neighbours_648121_tumour$Tumour <- rowSums(neighbours_648121_tumour==1)
neighbours_648121_tumour$Lymph <- rowSums(neighbours_648121_tumour==2)
neighbours_648121_tumour$Stroma <- rowSums(neighbours_648121_tumour==3)
neighbours_648121_tumour$Normal <- rowSums(neighbours_648121_tumour==4)

neighbours_648121_tumour_by_position <- data.frame(cbind(colSums(neighbours_648121_tumour==1)+colSums(neighbours_648121_tumour==4), colSums(neighbours_648121_tumour==2), colSums(neighbours_648121_tumour==3)))
colnames(neighbours_648121_tumour_by_position) <- c("Tumour", "Lymphocytes", "Stroma")
neighbours_648121_tumour_by_position$Position <- rownames(neighbours_648121_tumour_by_position)
neighbours_648121_tumour_by_position$Position[1:9] <- c("X01", "X02", "X03", "X04", "X05", "X06", "X07", "X08", "X09")
neighbours_648121_tumour_by_position_melted <- melt(neighbours_648121_tumour_by_position[1:50,])

ggplot(neighbours_648121_tumour_by_position_melted, aes(x=Position, y=value/nrow(neighbours_648121_tumour), group=variable, colour=variable))+
  geom_line(size=2)+
  theme_bw(base_size=18)+ 
  scale_color_discrete("Neighbouring Cell Type")+
  labs(title="Tumour cells neighbour cell type")+
  geom_hline(yintercept = c(0.409369812,	0.079632456,	0.510997732), colour=c("#F8766D", "#00BA38", "#619CFF" ), linetype="dashed")+ 
  theme(axis.text.x = element_text(angle = 90, hjust = 1))+
  labs(y="Percentage")


```
```{r}
neighbours_605012_tumour <- read_csv("~/Documents/OneDrive - University Of Cambridge/Documents/PhD/Neoadjuvant/Thomas/Cell_clustering/detail/clustering_detail_605012_tumour_cellids_nearest50.csv", col_names = FALSE)

neighbours_605012_tumour$ID <- paste("tumour",c(1:nrow(neighbours_605012_tumour)))
rownames(neighbours_605012_tumour) <- neighbours_605012_tumour$ID

neighbours_605012_tumour$Tumour <- rowSums(neighbours_605012_tumour==1)
neighbours_605012_tumour$Lymph <- rowSums(neighbours_605012_tumour==2)
neighbours_605012_tumour$Stroma <- rowSums(neighbours_605012_tumour==3)
neighbours_605012_tumour$Normal <- rowSums(neighbours_605012_tumour==4)

neighbours_605012_tumour_by_position <- data.frame(cbind(colSums(neighbours_605012_tumour==1)+colSums(neighbours_605012_tumour==4), colSums(neighbours_605012_tumour==2), colSums(neighbours_605012_tumour==3)))
colnames(neighbours_605012_tumour_by_position) <- c("Tumour", "Lymphocytes", "Stroma")
neighbours_605012_tumour_by_position$Position <- rownames(neighbours_605012_tumour_by_position)
neighbours_605012_tumour_by_position$Position[1:9] <- c("X01", "X02", "X03", "X04", "X05", "X06", "X07", "X08", "X09")
neighbours_605012_tumour_by_position_melted <- melt(neighbours_605012_tumour_by_position[1:50,])

ggplot(neighbours_605012_tumour_by_position_melted, aes(x=Position, y=value/nrow(neighbours_605012_tumour), group=variable, colour=variable))+
  geom_line(size=2)+
  theme_bw(base_size = 18)+ 
  scale_color_discrete("Neighbouring Cell Type")+
  labs(title="Tumour cells neighbour cell type")+
  geom_hline(yintercept = c(0.081928864,	0.03723156,	0.880839577), colour=c("#F8766D", "#00BA38", "#619CFF" ), linetype="dashed")+ 
  theme(axis.text.x = element_text(angle = 90, hjust = 1))+
  labs(y="Percentage")


```







Then let's look at distances

comparing 605012 and 648121

```{r}
two_slide_comparison <- read_csv("~/Documents/OneDrive - University Of Cambridge/Documents/PhD/Neoadjuvant/Thomas/two_slide_comparison.csv")

colnames(two_slide_comparison)

two_slide_summary <- two_slide_comparison[c(8, 16), 1:13]
two_slide_summary_melted <- melt(two_slide_summary, id=c("Slide_ID", "Cluster_Size"))



```

