---
title: "getting the cell id files"
output: pdf_document
---

Here we have 349 biopsy slides and we can see the identity of the nearest 50 neighbours for every cell.  

```{r}
library(readr)

#get the files
csvfolder <- "/Volumes/Wei/Cell_clustering/detail_newlist"
csv_list <- list.files(csvfolder, pattern = "_cellids", recursive = T, full.names = T)
#csv_list_lymphocytes <- csv_list[grep("lymphocyte", csv_list)]


all_files <- lapply(csv_list, read.csv)
names(all_files) <- sub("/Volumes/Wei/Cell_clustering/detail_newlist/clustering_detail_", "", csv_list)
names(all_files) <- sub("_cellids_nearest50.csv", "", names(all_files))

#now each file in the list of all_files has a name like "616851_lymphocyte" 

#we then create different list for lymphocytes, stroma and tumour (with normal included)
df_tumour <- all_files[c(grep("tumour", names(all_files)), grep("normal", names(all_files)))]
df_lymphocyte <- all_files[grep("lymphocyte", names(all_files))]
df_stroma <- all_files[grep("stroma", names(all_files))]

#now add slide ID and cell type to each dataframe for combining
for (i in 350:698) {
#    df_tumour[[i]]$ImageID <- sub("_tumour", "",df_tumour[[i]]$ImageID)
#   df_tumour[[i]]$ImageID <- sub("_normal", "", df_tumour[[i]]$ImageID)
#    df_tumour[[i]]$Base_celltype <- "Tumour"
 }
# 
# for (i in :349){
#    df_lymphocyte[[i]]$ImageID <- names(df_lymphocyte)[i]
#    df_lymphocyte[[i]]$Base_celltype <- "Lymphocyte"
#    df_stroma[[i]]$ImageID <- names(df_stroma)[i]
#    df_stroma[[i]]$Base_celltype <- "Stromal"
# }

saveRDS(df_tumour, file="df_tumour.RDS")
saveRDS(df_lymphocyte, file="df_lymphocyte.RDS")
saveRDS(df_stroma, file="df_stroma.RDS")

df_tumour <- readRDS("df_tumour.RDS")
```

We can calculate summary statistic.  We shall start with tumour cells.

```{r}
#tidy up column names
closest_colnames <- paste("C", c(1:50), sep = "")
image_id_list <- sub("_tumour", "", names(df_tumour)[1:349])
for (i in 1:length(df_tumour)){
  colnames(df_tumour[[i]]) <- c(closest_colnames, "ImageID", "Base_celltype") 
  }
#lapply(df_tumour, setNames, nm = c(closest_colnames, "ImageID", "Base_celltype"))
#colnames(df_tumour[[650]])
#now we check if the images are in the same order in tumours and normals before we combine them
#tumourlist <- names(df_tumour)[1:349]
#tumourlist <- sub("_tumour", "", tumourlist)
#normallist <- names(df_tumour)[350:698]
#normallist <- sub("_normal", "", normallist)
#mean(tumourlist==normallist)

#now that it is the same
df_alltumours <- list()
for (i in 1:349){
  df_alltumours[[i]] <- rbind(df_tumour[[i]], df_tumour[[i+349]])
  }

for (i in 1:349){
 for (j in 1:50){
  df_alltumours[[i]][,j]<-as.factor(df_alltumours[[i]][,j])
 }
}

```

we now practise it on a shortened version
```{r}
shortened <- readRDS("/Users/cope01/Documents/OneDrive - University Of Cambridge/Documents/PhD/Neoadjuvant/rscripts/shortened.RDS")

#combine all the normals and tumours in neighbouring cells ids
for (i in 1:349){
 for (j in 1:50){
  shortened[[i]][, j] <- as.factor(sub("4", "1", shortened[[i]][, j]))
 }
}

#now to work out the summary answer for each image
#create an empty dataframe
summary_tumour <- data.frame(matrix(0, nrow = 1047, ncol=52))
summary_tumour[,51] <- rep(c("1", "2", "3"), each=349)
colnames(summary_tumour)[51] <- "cell_type"
colnames(summary_tumour)[52] <- "ImageID"
colnames(summary_tumour)[1:50] <- closest_colnames

#it is important to note that we're no longer calculating on position 1, 2, 3 etc.  we are calculating a cumulative sum.  so now column c5 calculates the percentage of the 5 cells, not the 5th cell

for (i in 1:349){
    tmp_total <- table(factor(matrix(,ncol=3, nrow=1), levels=1:3))
  for (j in 1:50){
    tmp <- table(factor(shortened[[i]][,j], levels=1:3))
    tmp_total <- tmp+tmp_total
    summary_tumour[i,j] <- tmp_total[which(names(tmp_total)=="1")]/sum(tmp_total)
    summary_tumour[i+349,j] <- tmp_total[which(names(tmp_total)=="2")]/sum(tmp_total)
    summary_tumour[i+698,j] <- tmp_total[which(names(tmp_total)=="3")]/sum(tmp_total)
    summary_tumour[c(i, i+349, i+698), 52] <- names(shortened)[i]
  }
}



for (i in 1:349){
    tmp_total <- table(factor(matrix(,ncol=3, nrow=1), levels=1:3))
  for (j in 1:50){
    tmp <- table(factor(df_alltumours[[i]][,j], levels=1:3))
    tmp_total <- tmp+tmp_total
    summary_tumour[i,j] <- tmp_total[which(names(tmp_total)=="1")]/sum(tmp_total)
    summary_tumour[i+349,j] <- tmp_total[which(names(tmp_total)=="2")]/sum(tmp_total)
    summary_tumour[i+698,j] <- tmp_total[which(names(tmp_total)=="3")]/sum(tmp_total)
    summary_tumour[c(i, i+349, i+698), 52] <- names(df_alltumours)[i]
  }
}





```

collecting the metadata

```{r}
#load data
library(readr)
library(ggpubr)

#who has pCR
transneo_rcb_only <- read_csv("~/Documents/OneDrive - University Of Cambridge/Documents/PhD/Neoadjuvant/Pathology info/TransNeo_RCB.csv")

#transneo er her2 info etc
transneo_ErHer2 <- read_csv("~/Documents/OneDrive - University Of Cambridge/Documents/PhD/Neoadjuvant/Transneo_er_her2.csv")

transneo_rcb <- merge(transneo_rcb_only, transneo_ErHer2, by="Trial_ID")

#list of cases and image ID
combined_ffpe <- read_csv("~/Documents/OneDrive - University Of Cambridge/Documents/PhD/Neoadjuvant/Scanned Images/Combined_ffpe.csv")

#density
#catalogue_output <- read_csv("~/Documents/OneDrive - University Of Cambridge/Documents/PhD/Neoadjuvant/Image Data/NeoAdjuvant--20180808-145608.csv")

#then subset using only image ID that we have neighbourhood analysis for
biopsy_subset <- combined_ffpe[which(combined_ffpe$`Image ID` %in% image_id_list), ]

#now merge with the transneo_rcb data
subset_transneo <- merge(transneo_rcb, biopsy_subset, by="Trial_ID")
colnames(subset_transneo)[32] <- "ImageID"


#now combine the neighbourhood data with the metadata
tumour_with_meta <- merge(subset_transneo, summary_tumour, by="ImageID")
```

let's see if the neighbourhood info makes a difference

```{r}
colnames(tumour_with_meta)
tumour_with_meta <- tumour_with_meta[-which(is.na(tumour_with_meta$RCB)), ]

tumour_with_meta_sub <- tumour_with_meta[, c(1:34, 38, 43, 48, 53, 83, 84)]

tumour_with_meta_sub_melted <- melt(tumour_with_meta_sub, id=colnames(tumour_with_meta_sub)[c(1:33, 40)])
tumour_with_meta_melted <- melt(tumour_with_meta, id=colnames(tumour_with_meta)[c(1:33, 84)])
tumour_with_meta_melted$variable <- as.numeric(sub("C", "", tumour_with_meta_melted$variable))

#change the facet label from 1, 2, 3 to corresponding cell types
facet_labels <- c("1" = "Tumour Cells", "2" = "Lymphocytes", "3" = "Stromal Cells")



ggplot(tumour_with_meta_melted, aes(x=variable, y=value, colour=RCB)) +
  #geom_boxplot()+
  #geom_point()+
  geom_smooth(formula=y~x)+
  facet_wrap(~cell_type, labeller=labeller(cell_type = facet_labels))


ggplot(tumour_with_meta_melted, aes(x=variable, y=value, colour=cell_type))+
         geom_smooth(formula=y~x)+
         facet_wrap(~RCB)

tumour_wth_meta_melted_1to10 <- tumour_with_meta_melted[which(tumour_with_meta_melted$variable %in% c(1:10, 1)), ]

ggplot(tumour_wth_meta_melted_1to10, aes(x=variable, y=value, colour=RCB))+
         geom_smooth(formula=y~x)+
  facet_wrap(~cell_type)





```



```{r}
tumour_with_meta_sub_melted$RCB <- as.factor(tumour_with_meta_sub_melted$RCB)
tumour_with_meta_sub_melted$pCRornot <- ifelse(tumour_with_meta_sub_melted$RCB=="pCR", "pCR", "not_pCR")

ggplot(tumour_with_meta_sub_melted, aes(x=pCRornot,y=value))+
  geom_boxplot()+
  facet_grid(cell_type~variable)+
  theme_bw()+
  stat_compare_means()


```

now we use the real thing

```{r}
#get the file from the cluster
summary_tumour <- readRDS("/Users/cope01/Documents/OneDrive - University Of Cambridge/Documents/PhD/Neoadjuvant/rscripts/summary_tumour.RDS")
load("~/Documents/OneDrive - University Of Cambridge/Documents/PhD/Nanostring/R scripts/transneo_rcb_digipath.RDATA")
colnames(test_df_rcb)[1] <- "ImageID"
#colnames(summary_tumour)

combined_tumour <- merge(summary_tumour, test_df_rcb, by="ImageID")
normalised_combined_tumour <- combined_tumour

#now add the ratio (normalised) column
for (i in 1:nrow(normalised_combined_tumour)){
  for(j in 1:50){
  if (normalised_combined_tumour$cell_type[i]=="1") {
    normalised_combined_tumour[i,j+1] <- normalised_combined_tumour[i, j+1]/normalised_combined_tumour$fraction_cancer[i]
  } else if (normalised_combined_tumour$cell_type[i]=="2") {
    normalised_combined_tumour[i,j+1] <- normalised_combined_tumour[i, j+1]/normalised_combined_tumour$fraction_lymph[i]
  } else if (normalised_combined_tumour$cell_type[i]=="3") {
    normalised_combined_tumour[i,j+1] <- normalised_combined_tumour[i, j+1]/normalised_combined_tumour$fraction_stromal[i]
  }
}
}

```

see if we can plot it

```{r}
combined_tumour_melted <- melt(combined_tumour, id=colnames(combined_tumour)[c(1, 52:115)])
combined_tumour_melted$variable <- as.numeric(combined_tumour_melted$variable)
ggplot(combined_tumour_melted, aes(x=variable, y=value, colour=pCR)) +
  geom_smooth(formula = y~x)+
  facet_wrap(~cell_type, labeller=labeller(cell_type = facet_labels))+
  theme_bw()

combined_normalised_tumour_melted <- melt(normalised_combined_tumour, id=colnames(normalised_combined_tumour)[c(1, 52:115)])
combined_normalised_tumour_melted$variable <- as.numeric(combined_normalised_tumour_melted$variable)

ggplot(combined_normalised_tumour_melted, aes(x=variable, y=value)) +
  #geom_smooth(formula = y~x)+
  geom_point()+
  facet_wrap(~cell_type, labeller=labeller(cell_type = facet_labels))+
  theme_bw()+
  ylim(c(0, 20))

save(combined_tumour, file="combined_tumour.RData")

```


```{r}
#now just focus on tumour cells
tumour_cells_only <- combined_normalised_tumour_melted[which(combined_normalised_tumour_melted$cell_type=="1"), ]

ggplot(tumour_cells_only, aes(x=variable, y=value, colour=pCR)) +
  geom_point()+
  geom_line()+
  theme_bw()+
  facet_wrap(~RCB)

ggplot(data = tumour_cells_only) +
  geom_pointrange(mapping = aes(x = variable, y = value, colour=pCR),
                  stat = "summary",
                  fun.ymin = function(z) {quantile(z,0.25)},
                  fun.ymax = function(z) {quantile(z,0.75)},
                  fun.y = median)+
  theme_bw()



```

now we analyse similar things for lymphocytes

```{r}
csv_list_lymphocytes <- csv_list[grep("lymphocyte", csv_list)]
lymphocytes_files <- lapply(csv_list_lymphocytes, function(i) read.csv(i, header = FALSE))


names(lymphocytes_files) <- sub("/Volumes/Wei/Cell_clustering/detail_newlist/clustering_detail_", "", csv_list_lymphocytes)
names(lymphocytes_files) <- sub("_lymphocyte_cellids_nearest50.csv", "", names(lymphocytes_files))

saveRDS(lymphocytes_files, file="df_lymphocytes.RDS")
```

We can calculate summary statistic for lymphocytes
```{r}
#tidy up column names
closest_colnames <- paste("C", c(1:50), sep = "")
image_id_list <- names(lymphocytes_files)
for (i in 1:length(lymphocytes_files)){
  lymphocytes_files[[i]]$ImageID <- names(lymphocytes_files)[i]
  lymphocytes_files[[i]]$Base_celltype <- "Lymphocytes"
  colnames(lymphocytes_files[[i]]) <- c(closest_colnames, "ImageID", "Base_celltype") 
  }

#now change into factors
for (i in 1:349){
 for (j in 1:50){
  lymphocytes_files[[i]][,j]<-as.factor(lymphocytes_files[[i]][,j])
 }
}

#now the summary
summary_lymphocytes <- data.frame(matrix(0, nrow = 1047, ncol=52))
summary_lymphocytes[,51] <- rep(c("1", "2", "3"), each=349)
colnames(summary_lymphocytes)[51] <- "cell_type"
colnames(summary_lymphocytes)[52] <- "ImageID"
colnames(summary_lymphocytes)[1:50] <- closest_colnames
for (i in 1:349){
    tmp_total <- table(factor(matrix(,ncol=3, nrow=1), levels=1:3))
  for (j in 1:50){
    tmp <- table(factor(lymphocytes_files[[i]][,j], levels=1:3))
    tmp_total <- tmp+tmp_total
    summary_lymphocytes[i,j] <- tmp_total[which(names(tmp_total)=="1")]/sum(tmp_total)
    summary_lymphocytes[i+349,j] <- tmp_total[which(names(tmp_total)=="2")]/sum(tmp_total)
    summary_lymphocytes[i+698,j] <- tmp_total[which(names(tmp_total)=="3")]/sum(tmp_total)
    summary_lymphocytes[c(i, i+349, i+698), 52] <- names(lymphocytes_files)[i]
  }
}

saveRDS(summary_lymphocytes, file="summary_lymphocytes.RDS")

combined_lymphocytes <- merge(summary_lymphocytes, test_df_rcb, by="ImageID")
library(reshape2)
library(ggplot2)
combined_lymphocytes_melted <- melt(combined_lymphocytes, id=colnames(combined_lymphocytes)[c(1, 52:115)])
combined_lymphocytes_melted$variable <- as.numeric(combined_lymphocytes_melted$variable)
ggplot(combined_lymphocytes_melted, aes(x=variable, y=value, colour=pCR)) +
  geom_smooth(formula = y~x)+
  facet_wrap(~cell_type, labeller=labeller(cell_type = facet_labels))+
  theme_bw()

#now add the ratio (normalised) column
normalised_combined_lymphocytes <- combined_lymphocytes
for (i in 1:nrow(normalised_combined_lymphocytes)){
  for(j in 1:50){
  if (normalised_combined_lymphocytes$cell_type[i]=="1") {
    normalised_combined_lymphocytes[i,j+1] <- normalised_combined_lymphocytes[i, j+1]/normalised_combined_lymphocytes$fraction_cancer[i]
  } else if (normalised_combined_lymphocytes$cell_type[i]=="2") {
    normalised_combined_lymphocytes[i,j+1] <- normalised_combined_lymphocytes[i, j+1]/normalised_combined_lymphocytes$fraction_lymph[i]
  } else if (normalised_combined_lymphocytes$cell_type[i]=="3") {
    normalised_combined_lymphocytes[i,j+1] <- normalised_combined_lymphocytes[i, j+1]/normalised_combined_lymphocytes$fraction_stromal[i]
  }
}
}



combined_normalised_lymphocytes_melted <- melt(normalised_combined_lymphocytes, id=colnames(normalised_combined_lymphocytes)[c(1, 52:115)])
combined_normalised_lymphocytes_melted$variable <- as.numeric(combined_normalised_lymphocytes_melted$variable)
ggplot(combined_normalised_lymphocytes_melted, aes(x=variable, y=value, colour=pCR)) +
  geom_smooth(formula = y~x)+
  facet_wrap(~cell_type, labeller=labeller(cell_type = facet_labels))+
  theme_bw()+
  ggtitle("Normalised for cell type fraction")

save(combined_lymphocytes, file="combined_lymphocytes.RData")

```

now we do stromal cells

```{r}
csv_list_stromal <- csv_list[grep("stroma", csv_list)]
stroma_files <- lapply(csv_list_stromal, function(i) read.csv(i, header = FALSE))

names(stroma_files) <- sub("/Volumes/Wei/Cell_clustering/detail_newlist/clustering_detail_", "", csv_list_stromal)
names(stroma_files) <- sub("_stroma_cellids_nearest50.csv", "", names(stroma_files))

saveRDS(stroma_files, file="df_stroma.RDS")
```

```{r}
#tidy up column names
closest_colnames <- paste("C", c(1:50), sep = "")
image_id_list <- names(stroma_files)
for (i in 1:length(stroma_files)){
  stroma_files[[i]]$ImageID <- names(stroma_files)[i]
  stroma_files[[i]]$Base_celltype <- "Stroma"
  colnames(stroma_files[[i]]) <- c(closest_colnames, "ImageID", "Base_celltype") 
  }

#now change into factors
for (i in 1:349){
 for (j in 1:50){
  stroma_files[[i]][,j]<-as.factor(stroma_files[[i]][,j])
 }
}

#now the summary
summary_stroma <- data.frame(matrix(0, nrow = 1047, ncol=52))
summary_stroma[,51] <- rep(c("1", "2", "3"), each=349)
colnames(summary_stroma)[51] <- "cell_type"
colnames(summary_stroma)[52] <- "ImageID"
colnames(summary_stroma)[1:50] <- closest_colnames
for (i in 1:349){
    tmp_total <- table(factor(matrix(,ncol=3, nrow=1), levels=1:3))
  for (j in 1:50){
    tmp <- table(factor(stroma_files[[i]][,j], levels=1:3))
    tmp_total <- tmp+tmp_total
    summary_stroma[i,j] <- tmp_total[which(names(tmp_total)=="1")]/sum(tmp_total)
    summary_stroma[i+349,j] <- tmp_total[which(names(tmp_total)=="2")]/sum(tmp_total)
    summary_stroma[i+698,j] <- tmp_total[which(names(tmp_total)=="3")]/sum(tmp_total)
    summary_stroma[c(i, i+349, i+698), 52] <- names(stroma_files)[i]
  }
}

saveRDS(summary_stroma, file="summary_stroma.RDS")

combined_stroma <- merge(summary_stroma, test_df_rcb, by="ImageID")
library(reshape2)
library(ggplot2)
combined_stroma_melted <- melt(combined_stroma, id=colnames(combined_stroma)[c(1, 52:115)])
combined_stroma_melted$variable <- as.numeric(combined_stroma_melted$variable)
ggplot(combined_stroma_melted, aes(x=variable, y=value, colour=pCR)) +
  geom_smooth(formula = y~x)+
  facet_wrap(~cell_type, labeller=labeller(cell_type = facet_labels))+
  theme_bw()

#now add the ratio (normalised) column
normalised_combined_stroma <- combined_stroma
for (i in 1:nrow(normalised_combined_stroma)){
  for(j in 1:50){
  if (normalised_combined_stroma$cell_type[i]=="1") {
    normalised_combined_stroma[i,j+1] <- normalised_combined_stroma[i, j+1]/normalised_combined_stroma$fraction_cancer[i]
  } else if (normalised_combined_stroma$cell_type[i]=="2") {
    normalised_combined_stroma[i,j+1] <- normalised_combined_stroma[i, j+1]/normalised_combined_stroma$fraction_lymph[i]
  } else if (normalised_combined_stroma$cell_type[i]=="3") {
    normalised_combined_stroma[i,j+1] <- normalised_combined_stroma[i, j+1]/normalised_combined_stroma$fraction_stromal[i]
  }
}
}



combined_normalised_stroma_melted <- melt(normalised_combined_stroma, id=colnames(normalised_combined_stroma)[c(1, 52:115)])
combined_normalised_stroma_melted$variable <- as.numeric(combined_normalised_stroma_melted$variable)
ggplot(combined_normalised_stroma_melted, aes(x=variable, y=value, colour=pCR)) +
  geom_smooth(formula = y~x)+
  facet_wrap(~cell_type, labeller=labeller(cell_type = facet_labels))+
  theme_bw()

save(combined_stroma, file="combined_stroma.RData")

```